Persistence in functional topology and a correction to a theorem of Morse

### 2021-07-15 vietoris.tex
Before meeting on Friday mid July

\section{Vietoris homology} \label{s:vietoris}

In his work on functional topology, Morse uses the Vietoris homology for metric spaces as introduced in \cite{MR1512371}. So far, we have considered \v{C}ech homology instead, applying a topological rather than metric construction.
We will recall the definition of Vietoris homology and recall that these two constructions yield the same result for compact metric spaces.

\v{C}ech homology of a pair of spaces $(X_{1}, X_{2})$ was defined as the inverse limit over the homology of nerves of covers of $(X_{1}, X_{2})$.
As an alternative to the nerve construction, for a cover $\alpha = (\alpha_{1}, \alpha_{2})$ one can define $\Vietoris(\alpha)$ as the simplicial pair $\big(\Vietoris (\alpha_{1}), \Vietoris (\alpha_{2})\big)$, where 
\[
\Vietoris (\alpha_{i}) = \left\{ \sigma \subseteq X_{i} \mid \sigma \text{ is finite and there is } U \in \alpha_{i} \text{ with } \sigma \in U \right\}.
\]
This construction is dual in the sense of Dowker's Theorem \cite{Dowker.1952} to the nerve construction.
As a consequence, we have that
$
H_{*} (\Nrv (\alpha); G) \cong H_{*} (\Vietoris (\alpha); G)
$
for any coefficient group $G$.
This isomorphism is natural with respect to refinement and compatible with the boundary operators, so that we get an alternative description of \v{C}ech homology as 
%the inverse limit over the homology of Vietoris complexes of covers.
\[
\lim_{\alpha \in \Cov (X_{1}, X_{2})} H_{*} (\Vietoris (\alpha); G) \, \cong\, \CH (X_{1}, X_{2}; G)
\]
This is still not exactly the construction of Vietoris homology of a pair of metric spaces $(X_{1}, X_{2})$ as presented in \cite{MR1512371}, which in modern notation is the limit
\begin{equation*}
    \lim_{\alpha \in \Balls(X_{1}, X_{2})} H_{*} (\Vietoris (\alpha); G),
\end{equation*}
where 
\[
\Balls (X_{1}, X_{2}) = \left\{ \big( ( B_{\delta} (x) )_{x \in X_{1}} , ( B_{\delta} (x) )_{x \in X_{2}} \big) \mid \delta > 0 \right\}
%\left\{ \big(B_{\delta}(x),\ B_{\delta}(y) \big) \right\}_{(x,y) \in X_{1} \times X_2} 
\subseteq \Cov (X_{1}, X_{2}).
\]
But if the metric pair $(X_{1}, X_{2})$ is compact, then $\Balls (X_{1}, X_{2})$ is coinitial in $\Cov (X_{1}, X_{2})$, that is to say, they define the same limit.
Thus, we have natural isomorphisms
\[
\lim_{\alpha \in \Balls(X_{1}, X_{2})} H_{*} (\Vietoris (\alpha); G) \ \cong \,
%\lim_{\alpha \in \Cov (X_{1}, X_{2})} H_{*} (\Vietoris (\alpha); G) \, \cong\, 
\CH (X_{1}, X_{2}; G),
\]
so in Morse's setting the metric Vietoris homology theory he uses is indeed the same as the \v{C}ech homology we considered earlier on.

>>>>

\section{Vietoris and \v{C}ech homology} \label{s:vietoris}

As we have mentioned before, Morse's original treatment of functional topology relied on him studying the Vietoris homology \cite{MR1512371} of the sublevel set filtration induced by a function on a metric space. 
For ease of reference, we have instead talked about \v{C}ech homology, which will now be justified by recalling that these two homology constructions agree for compact metric spaces.

First, we recall the definition of \v{C}ech homology as presented for example by \citet[Section IX--X]{MR0050886}.
Let $X$ be a topological space and let $\Cov(X)$ be the set of all open covers of $X$ ordered by the refinement relation. 
Recall that for an open cover $\alpha \in \Cov(X)$ its \emph{nerve} $\Nrv(\alpha)$ is defined as the simplicial complex
\[
\Nrv(\alpha) =
\big\{ \beta \subseteq \alpha \mid \beta \text{ is finite and } \textstyle{\bigcap_{U \in \beta}} \, U \neq \emptyset \big\}.
\]
The nerve construction defines a functor from the poset $\Cov(X_{1},X_{2})$ regarded as a category to the category of simplicial complexes. 
The \emph{\v{C}ech homology with coefficients in $\mathbb{F}$} of $X$ is defined as
\[
\CH_{*}(X; \mathbb{F}) \ =
\lim_{\alpha \in \Cov(X)} H_{*}(\Nrv(\alpha); \mathbb{F}),
\]
which extends to a homotopy invariant functor and which admits boundary operators in a natural way.

As an alternative to the nerve construction, for a cover $\alpha \in \Cov(X)$ one can define $\Vietoris(\alpha)$ as the simplicial complex
\[
\Vietoris (\alpha) = \left\{ \sigma \subseteq X \mid \sigma \text{ is finite and } \sigma \in U \text{ for some } U \in \alpha  \right\},
\]
which again yields a functor from $\Cov(X)$ to simplicial complexes.
This construction is dual in the sense of Dowker's Theorem \cite{Dowker.1952} to the nerve construction.
As a consequence, we have that
$
H_{*} (\Nrv (\alpha); \mathbb{F}) \cong H_{*} (\Vietoris (\alpha); \mathbb{F}).
$
This isomorphism is natural with respect to refinement and compatible with the boundary operators, so that we get an alternative description of \v{C}ech homology as 
%the inverse limit over the homology of Vietoris complexes of covers.
\[
\CH (X; \mathbb{F}) \ =
\lim_{\alpha \in \Cov (X)} H_{*} (\Vietoris (\alpha); \mathbb{F}).
\]

If $X$ is a metric space, this is still not exactly the same as the construction of Vietoris homology as presented in \cite{MR1512371} and as used by Morse, which in modern notation is the limit
\begin{equation*}
    \lim_{\alpha \in \Balls(X)} H_{*} (\Vietoris (\alpha); \mathbb{F}),
\end{equation*}
where 
\[
\Balls (X) = \left\{ ( B_{\delta} (x) )_{x \in X} \mid \delta > 0 \right\}
\subseteq \Cov (X).
\]
However, if the metric space $X$ is compact, then $\Balls (X)$ is coinitial in $\Cov (X)$, that is to say, they define the same limit.
Thus, we have a natural isomorphism
\[
\CH_{*} (X; \mathbb{F}) \ \cong \,
\lim_{\alpha \in \Balls(X)} H_{*} (\Vietoris (\alpha); \mathbb{F})
\]
for a compact metric space $X$ and so the metric Vietoris homology theory employed in Morse's setting is indeed canonically isomorphic to the \v{C}ech homology we considered earlier on.
